---
title: "fertplan introduction"
output: github_document
---

## Description

The goal of the package is to provide the necessary computation algorithm to perform a fertilization plan for the fields of a farm. It heavily follows the agronomic guidelines for integrated agriculture, issued by [Lazio Region](https://www.regione.lazio.it/rl_agricoltura/?vw=contenutidettaglio&id=164 "Web-site of Lazio Region (in italian)"), a public administration in Italy.
Fertilization plans in the Lazio region territory have to follow these agronomic guidelines with specific attention to [attachment no. 2](http://www.regione.lazio.it/binary/rl_main/tbl_documenti/AGC_DD_G01782_24_02_2020_Allegato1.pdf "PDF file of the Attachment 2 of the guidelines") (from now on *the guidelines*).

The package provides a set of functions to compute the components of the supply/demand for Nitrogen, Phosphorus $P_2O_5$, and Potassium $K_2O$ nutrients to field crops.


## Installation

`fertplan` is currently in active development and it not yet on CRAN, it may be installed from this GitHub repository though:

```{r, eval = FALSE}
# Install devtools package if not yet present in R library
# install.packages("devtools")

devtools::install_github("mbask/fertplan")
```


## Usage

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
library(data.table)
library(fertplan)
```

This document will walk you through a simulation of a real fertilization plan for nitrogen nutrient. Both `fertplan` and this document depend on package `data.table` but its usage is not in any way mandatory.

```{r, results='asis', echo = FALSE}
res <- knitr::knit_child('vignettes/n_fertplan.Rmd', quiet = TRUE)
cat(res, sep = '\n')
```


Please check out the package vignettes for:

  * Nitrogen fertilization plan
  * Phosphorus fertilization plan
  * Potassium fertilization plan
