---
title: "Introduction to `fertplan` R package"
output: github_document
references:
- id: guidelines2020
  title: Parte agronomica, norme generali, Disciplinare di Produzione Integrata della Regione Lazio - SQNPI
  author:
  - family: Assessorato Agricoltura, Promozione della Filiera e della Cultura del Cibo, Ambiente e Risorse Naturali
  URL: 'http://www.regione.lazio.it/rl_agricoltura/?vw=documentazioneDettaglio&id=52065'
  publisher: Regione Lazio
  type: report
  issued:
    year: 2020
---

## Description

The goal of the package is to provide the necessary computation algorithm to perform a fertilization plan for the fields of a farm. It heavily follows the agronomic guidelines for integrated agriculture, issued by [Lazio Region](https://www.regione.lazio.it/rl_agricoltura/?vw=contenutidettaglio&id=164 "Web-site of Lazio Region (in italian)"), a public administration in Italy.
Fertilization plans in the Lazio region territory have to follow these agronomic guidelines with specific attention to [attachment no. 2](http://www.regione.lazio.it/binary/rl_main/tbl_documenti/AGC_DD_G01782_24_02_2020_Allegato1.pdf "PDF file of the Attachment 2 of the guidelines") [@guidelines2020].

The package provides a set of functions to compute the components of the supply/demand for Nitrogen, Phosphorus $P_2O_5$, and Potassium $K_2O$ nutrients to field crops.


## Installation

`fertplan` is currently in active development and not yet on CRAN, it may be installed from this GitHub repository though:

```{r, eval = FALSE}
# Install devtools package if not yet present in R library
# install.packages("devtools")

devtools::install_github("mbask/fertplan")
```


## Usage

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
library(data.table)
library(fertplan)
```

Please check out available package vignettes for:

  * Nitrogen fertilization plan
  * Phosphorus fertilization plan
  * Potassium fertilization plan


```{r, results='asis', echo = FALSE}
res <- knitr::knit_child('vignettes/n_fertplan.Rmd', quiet = TRUE)
cat(res, sep = '\n')
```


