---
title: "Potassium `fert`ilization `plan` for erbaceous and tree crops"
author: "Marco Bascietto"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Potassium plan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
references:
- id: guidelines2020
  title: Parte agronomica, norme generali, Disciplinare di Produzione Integrata della Regione Lazio - SQNPI
  author:
  - family: Assessorato Agricoltura, Promozione della Filiera e della Cultura del Cibo, Ambiente e Risorse Naturali
  URL: 'http://www.regione.lazio.it/rl_agricoltura/?vw=documentazioneDettaglio&id=52065'
  publisher: Regione Lazio
  type: report
  issued:
    year: 2020
---

```{r k_setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>")
library(data.table)
library(fertplan)
```

This document will walk you through a simulation of a real fertilization plan for potassium nutrient ($K_2O_5$). Both `fertplan` and this document depend on package `data.table` but its usage is not in any way mandatory.


## Potassium fertilization plan

This estimation of Phospohorus fertilization concentrations for erbaceous and tree crops will strictly follow the indications formulated in the regulation drawn up by the Italian Region of Lazio [@guidelines2020], hereafter the *guidelines*.

Potassium fertilization concentration in kg/ha is estimated as the net resultant of a K balance between the potassium pool available for the crops and the potassium losses. The K balance involves 3 flow components. Similarly to N plans, flows that increase K availability to the crop are > 0 (positive sign). Flows that deplete soil K pool or PN availability for the crop are < 0 (negative sign). 

The N flow components include:

  * **$f_{K,e}$** Crop demand for potassium on the basis of its expected yield
  * **$f_{K,f}$** Potassium concentration currently in the soil due to its fertility. Three multiplicative sub-components contribute to this flow: soil depth, apparent soil density, and offset from average "normal" soil K concentration.
  * **$f_{K,g}$** Potassium immobilized due to clay content
  * **$f_{K,h}$** Potassium lost due to leaching as a function of soil texture (specifically clay content)

The final potassium balance is computed as:
$$B_K = f_{K,e} + f_{K,f} \cdot f_{K,g} + f_{K,h}$$


### First step: load soil analyses

Let's begin with some minimal data from soil physical and chemical analyses on a few sampling points in the field.

```{r}
soil_dt <-  data.table::data.table(
  id      = factor(x = c("11", "20", "13", "12", "17")), 
  K_ppm   = c(324, 285, 398, 297, 476), 
  Clay_pc = c(37, 37, 38, 40, 36))
knitr::kable(soil_dt)
```


The table shows the soil chemical and physical status before the planned crop sowing. The soil analyses elements that will be fed the potassium balance estimation are:

  * *K_ppm*, Total potassium content in ppm (eg. mg/kg)
  * *Clay_pc*, Clay content in %

The *id* feature is not relevant to the balance estimation.


### Second step: variable configuration

A few environmental and crop-related variables need to be set. 
Some variables need to match those set out in the *guidelines* tables, while a few others have to be derived from external sources. 

Matching-variables are:

  * **Crop**, this is the name of the crop to be sown and will be used to lookup the its Potassium demand in table 15.2 (page 63) of the *guidelines* to contribute to **$f_{K,e}$** component. The name must match one of the following crop names available. Partial matching is allowed, provided that the partial string is unique among crop names. The allowed crop names is:
  ```{r, echo = FALSE}
knitr::kable(unique(fertplan:::tables_l$all_01_dt[, crop]))
```

  * **Texture**, soil texture, one of `r levels(fertplan:::tables_l$tab_01_wdt[, soil_texture])`. Soil texture enters in a **$f_{K,f}$** flow of the balance.


Environmental and crop-related variables include:

  * **Expected yield**, it contributes to **$f_{K,e}$** component, unit of measure kg/ha. It can be estimated from statistical [estimates](http://dati.istat.it "ISTAT web site") of crop areas and yields at province, regional, or national level. As an example, wheat expected yield is 2,900 kg/ha in the province of Rome, based on 2019 Istat estimates.

  * **Soil depth**, depth of soil tillage practise, in $cm$. This is usually 30 or 40 cm, for shallow tillage (the former) or deep tillage (the latter). A soil depth coefficient measured as depth / 10 enters the **$f_{K,f}$** flow component as a multiplicative coefficient (eg 3 for shallow tillage or 4 for deep tillage).

  
Let's now set the variables values and bind them to the soil analysis table. Let's suppose the values are constant among all soil samples, as it may be the case when all sampling points come from a uniform field that will be sown with the same crop:

  ```{r}
soil_dt[
  , `:=` (
    crop                 = "Grano duro (granella)",
    expected_yield_kg_ha = 2900,
    texture              = "Loam",
    soil_depth_cm        = 30)]
knitr::kable(soil_dt)
```



### Third step: estimate the components of K balance

Let's estimate the **$f_{K,A|B|C}$** components:

```{r}
soil_dt[
  , `:=` (
    E_K_kg_ha              = fertplan::A_crop_demand(
      crop_abs       = fertplan::rem_K_coef_of(crop) / 100,
      crop_exp_yield = expected_yield_kg_ha),
    F_K_kg_ha              = fertplan::F_K_in_soil(
      k_ppm         = K_ppm, 
      soil_texture  = texture, 
      soil_depth_cm = soil_depth_cm),
    G_K_kg_ha              = fertplan::G_K_immob_by_clay(Clay_pc),
    H_K_kg_ha              = fertplan::H_K_leaching(Clay_pc)  )]
```

All components were estimated. Remember that positive values are demand pools of K in soil or K flows leaving the field; negative values are current K pools in the soils that are available for assimilation to the crop or that will be available during the time-frame of crop growth.

```{r}
fertzl_cols <- grep(
  pattern = "^[A-Z]_K_kg_ha$", 
  x       = colnames(soil_dt), 
  value   = TRUE)
id_fertzl_cols <- c("id", fertzl_cols)
fertzl_dt <- soil_dt[, ..id_fertzl_cols]
knitr::kable(fertzl_dt)
```


### Fourth step: estimate K demand

We are finally arrived to the last step of assembling all components of the K balance. Let's compute the balance following the previous equation:
  
  ```{r}
fertzl_dt[, k_demand_kg_ha := E_K_kg_ha + F_K_kg_ha * G_K_kg_ha + H_K_kg_ha]
knitr::kable(fertzl_dt[, c("id", "k_demand_kg_ha")])
```

All sampling points end up needing a supply of Potassium of `r fertzl_dt[, mean(k_demand_kg_ha)]` kg/ha on average, that is there is a heavy surplus of P nutrient in the soil and fertilization is not needed.

That's it as far as potassium fetilization plan is concerned.

## References

